\begin{abstract}
Self-Consistency (SC) improves large language model (LLM) reasoning via majority voting over chain-of-thought (CoT) traces.
A natural extension is to extract formal constraints and aggregate them with a symbolic solver; but does it work?
Under zero-shot extraction, we test this through Symbolic Invariant Consensus Aggregation (SICA), which extracts first-order logic constraints and solves weighted partial MAX-SAT.
Across six models (7B--32B), four benchmarks, and over 1{,}200 instances, SICA reduces to majority voting: no condition shows significant improvement after FDR control; 4 of 11 conditions pass TOST equivalence tests ($\pm 2$\,pp margin); the remaining 7 are underpowered (MDE $4.6$--$4.8$\,pp) but show consistent directional evidence ($|\Delta| \leq 4.41$\,pp).
We diagnose this as \emph{symbolic confirmation bias}: under zero-shot self-extraction, constraints systematically confirm their source trace's answer, with bias ratios $1.66$--$4.97\times$ across nine models up to frontier scale.
The mechanism is formalized through effective ensemble size $K_\text{eff}$ from inter-trace agreement $\kappa$: correlated traces collapse nominal $K{=}12$ to $K_\text{eff}$ as low as 1.20.
Twenty-seven training-free remediation strategies fail to break this bottleneck.
As preliminary evidence (exploratory; $n{=}6$ generator--verifier pairs), a cross-architecture encoder-only verifier escapes the bias ($+6.67$\,pp, $p < 0.001$), suggesting that architectural independence can restore corrective signal.
Three zero-cost diagnostics predict failure modes from a pilot run\footnote{Code and data are available at \url{https://github.com/anonymous-nlp-2026/symb_invariant_consensus}.}.
\end{abstract}
