¥section{Method}
¥label{sec:method}
¥label{sec:pipeline}
¥label{sec:condorcet}
¥label{sec:effective-k}

Self-Consistency (SC) ¥citep{wang2023sc} samples $K$ reasoning traces and selects the majority answer. Under correlated traces (Fleiss' $¥kappa$ ¥citep{fleiss1971kappa}), the effective ensemble size $K_¥text{eff}$ is
¥begin{equation}
¥label{eq:effk}
K_{¥text{eff}} = K ¥,/¥, [1 + (K{-}1) ¥cdot ¥kappa]
¥end{equation}
(following ¥citet{ladha1992correlated}). Intuitively, $K_¥text{eff}$ measures how many effectively independent votes $K$ correlated traces contribute; e.g., $K{=}12$ traces with $¥kappa{=}0.69$ yield $K_¥text{eff} ¥approx 1.4$, meaning twelve correlated traces carry no more independent information than 1.4 uncorrelated ones. This motivates constraint-based aggregation.

Symbolic Invariant Consensus Aggregation (SICA) extends SC by extracting formal constraints from traces and selecting the answer best supported by the satisfiable subset (Figure‾¥ref{fig:pipeline}).

¥begin{table*}[t]
¥centering
¥small
¥caption{SICA vs.¥ SC across all non-degenerate self-extraction conditions (sorted by SC). Two conditions show significant degradation after Benjamini--Hochberg (BH) correction ($m{=}6$ confirmatory; remaining exploratory); no condition shows a significant gain. $¥Delta$ = SICA $-$ SC (pp). PW: ProofWriter. $^b$gold ICL; ¥textbf{bold}: $p < 0.05$; {*}¥,BH adj.¥ $p = 0.042$; $^{3s}$/$^{4s}$¥,multi-seed aggregates; $^¥dagger$¥,4-seed replication (Table‾¥ref{tab:multi-seed}); $^¥ddagger$¥,pilot ($n{¥leq}30$). Unmarked: single-seed exploratory (seed 123, not selected post hoc). Degenerate conditions and power analysis in Appendices‾¥ref{app:oracle} and‾¥ref{app:power}.}
¥label{tab:main_results}
¥begin{tabular}{@{}llrrrrl@{}}
¥toprule
Model & Domain & $n$ & SC (¥%) & SICA (¥%) & $¥Delta$ (pp) & $p$ ¥¥
¥midrule
¥multicolumn{7}{l}{¥textit{Type C: error amplification (SC $¥leq$ 40¥%)}} ¥¥
Mistral-7B & PW-D5 & 600 & 39.33 & 36.50 & $¥mathbf{-2.83}$ & 0.014{*} ¥¥
¥midrule
¥multicolumn{7}{l}{¥textit{Intermediate SC (SC $¥sim$50--60¥%, Mistral-7B only)}} ¥¥
Mistral-7B & LogiQA & 200 & 50.00 & 51.50 & $+$1.50 & 0.581 ¥¥
Mistral-7B & FOLIO (s123) & 204 & 54.41 & 57.35 & $+$2.94 & 0.070 ¥¥
Mistral-7B$^{3s}$ & FOLIO & 204 & 56.53 & 56.70 & $+$0.16 & 1.000 ¥¥
Mistral-7B & FOLIO (s456) & 204 & 55.88 & 55.88 & 0.00 & 1.000 ¥¥
Mistral-7B & FOLIO (s789) & 204 & 59.31 & 56.86 & $-$2.45 & 0.267 ¥¥
¥midrule
¥multicolumn{7}{l}{¥textit{Type A: confirmation bias (SC $¥geq$ 60¥%)}} ¥¥
LLaMA-3.1-8B & FOLIO & 204 & 66.18 & 61.76 & $-$4.41 & 0.093 ¥¥
LLaMA-3.1-8B & StrategyQA & 200 & 72.00 & 74.00 & $+$2.00 & 0.424 ¥¥
¥addlinespace
Qwen2.5-14B & FOLIO (ICL)$^b$ & 199 & 74.37 & 74.87 & $+$0.50 & 1.000 ¥¥
Qwen2.5-14B & PW-D5 & 600 & 74.00 & 71.50 & $¥mathbf{-2.50}$ & 0.011{*} ¥¥
Qwen2.5-14B$^{4s}$ & FOLIO & 204 & 75.01 & 73.78 & $-$1.24 & 1.000 ¥¥
Qwen2.5-7B & FOLIO & 204 & 75.98 & 76.47 & $+$0.49 & 1.000 ¥¥
Qwen2.5-14B & LogiQA & 200 & 79.00 & 78.00 & $-$1.00 & 0.625 ¥¥
¥addlinespace
Qwen3-14B & FOLIO & 204 & 83.33 & 84.80 & $+$1.47 & 1.000 ¥¥
Qwen3-14B$^¥dagger$ & PW-D5 & 600 & 85.83 & 85.67 & $-$0.17 & 1.000 ¥¥
Qwen3-14B (think) & FOLIO & 204 & 86.27 & 86.76 & $+$0.49 & 1.000 ¥¥
¥bottomrule
¥end{tabular}
¥end{table*}

¥paragraph{Trace generation.} Given question $q$, an LLM generates $K$ traces $¥{t_1, ¥ldots, t_K¥}$ at temperature $T = 0.7$. Each trace $t_i$ yields answer $a_i$.

¥looseness=-1
¥paragraph{Constraint extraction.}¥label{sec:constraint-extraction} The same LLM extracts logical constraints from each trace at $T_{¥text{ext}} = 0.3$. The extraction prompt requests JSON output with constraint type, expression, and Z3-compatible formula (Appendix‾¥ref{app:prompts}); in practice, 99.0¥% of extracted constraints are propositional at the Z3 level, with only 1.0¥% using true quantifiers. The ¥emph{extraction rate} is the fraction of traces yielding at least one parseable constraint.

¥paragraph{Deduplication and MAX-SAT.} Constraints are deduplicated by syntactic equivalence (global average deduplication ratio: 46.0¥%); each unique constraint $c_m$ receives weight $w_m$ equal to the number of source traces. The objective finds the largest-weight satisfiable subset:
¥begin{equation}
¥label{eq:maxsat}
¥mathcal{C}^* = ¥arg¥max_{¥mathcal{S} ¥subseteq ¥mathcal{C}} ¥sum_{c ¥in ¥mathcal{S}} w_c ¥quad ¥text{s.t.} ¥quad ¥mathcal{S} ¥text{ is satisfiable.}
¥end{equation}
Z3 ¥citep{demoura2008z3} solves this with a 10-second timeout (Algorithm‾¥ref{alg:maxsat}; Appendix‾¥ref{app:z3-pseudocode}). Answer assertion constraints ($¥texttt{answer¥_a} == ¥texttt{True}$, weight 1) ensure every candidate receives a nonzero base score; SICA is thus a strict generalization of SC.

¥paragraph{Scoring.} Let $¥mathcal{C}^{¥setminus} = ¥mathcal{C} ¥setminus ¥mathcal{C}^*$. Each candidate $a$ is scored by:
¥begin{equation}
¥label{eq:score}
¥text{score}(a) = ¥sum_{c ¥in ¥mathcal{C}^*} w_c ¥cdot ¥sigma(c, a) - ¥alpha ¥sum_{c ¥in ¥mathcal{C}^{¥setminus}} w_c ¥cdot ¥sigma(c, a),
¥end{equation}
where $¥sigma(c, a) = 1$ if $c$ was extracted from any trace producing answer $a$, and $¥alpha = 0.5$. The selected answer is $a^* = ¥arg¥max_{a} ¥text{score}(a)$.

¥paragraph{Degeneration condition.} If self-extracted constraints are coupled to their source answer, MAX-SAT retains nearly all ($|¥mathcal{C}^{¥setminus}| ¥approx 0$), and scoring degenerates to:
¥begin{equation}
¥label{eq:degenerate}
¥text{score}(a) ¥approx ¥sum_{¥substack{c ¥in ¥mathcal{C}^* ¥¥ ¥text{source}(c) ¥to a}} w_c,
¥end{equation}
equivalent to majority voting. We formalize this as Observation‾1 (¥S¥ref{sec:confirmation-bias}) and show empirically that all three conditions hold in practice.

¥paragraph{The $C_0$ diagnostic.} The multi-candidate rate $C_0 = |¥{q : |¥mathcal{A}(q)| > 1¥}| / |¥mathcal{Q}|$ is a lightweight diagnostic computable from pilot traces: $C_0 < 5¥%$ means SICA reduces to SC. $¥alpha = 0.5$ is not critical (${¥pm}0.49$¥,pp across $¥alpha ¥in [0, 1]$).

¥paragraph{Bias ratio.} We quantify extraction bias via the ¥emph{bias ratio} (BR):
¥begin{equation}¥label{eq:mirage}
¥text{BR}(M,T) = ¥frac{P(c ¥text{ supports } a_M ¥mid a_M ¥neq y)}{P(c ¥text{ supports } y ¥mid a_M ¥neq y)},
¥end{equation}
where $c$ is a self-extracted constraint, $a_M$ the model's majority answer, and $y$ the ground truth. BR quantifies the degree of extraction degeneration under source-attribution scoring (Observation‾1): values above 1 reflect the trace-count base rate inherent to source-attributed constraints. $¥text{BR} = 1$ indicates no coupling beyond chance. We validate this metric empirically in ¥S¥ref{sec:confirmation-bias}.
