\section{Escaping the Bottleneck}
\label{sec:solutions}
\label{sec:cross-arch}

\paragraph{Motivation.}
\looseness=-1 No training-free self-extraction method achieves independence from the model's distributional prior (\S\ref{sec:remediation}). The DPI bound implies that any escape must break the self-extraction dependency: $I(C;Y) \leq I(T;Y)$ binds whenever constraints are deterministic functions of the generating model's traces. We test whether a structurally independent verifier, sharing neither architecture nor training data with the generator, provides corrective signal that escapes this bound.

\paragraph{Method.}
\looseness=-1 We use DeBERTa-large \citep{he2021deberta} (406M, encoder-only), fine-tuned on MultiNLI \citep{williams2018mnli}, to classify (premise, conclusion) pairs as entailment, contradiction, or neutral. In combo mode, each NLI-supporting vote counts $w$ times alongside SC's $K$ votes. We evaluate on Mistral-7B FOLIO ($n = 204$) and ProofWriter depth-5 ($n = 600$).

\paragraph{Zero-shot results.}
\looseness=-1 DeBERTa NLI alone achieves 49.51\% on FOLIO and 55.00\% on ProofWriter. Combo $w{=}3$ yields $+6.67$\,pp on ProofWriter for Mistral-7B ($p < 0.001$, significant after BH correction) and $+2.94$\,pp on FOLIO (n.s.); all generators with SC $\geq 72\%$ show no significant improvement ($\Delta \leq +1.47$\,pp). Same-model SICA \emph{degrades} ProofWriter ($-2.83$\,pp; Table~\ref{tab:main_results}), illustrating that corrective signal must originate outside the model's distribution.

\paragraph{Capability gap.}
\looseness=-1 Across 108 verifier-generator-weight combinations spanning three NLI architectures, four generators, and three domains, gain decreases monotonically with generator accuracy---a \emph{capability-gap pattern}. Of 72 above-verifier conditions, none shows a significant improvement; all significant gains concentrate below the verifier's standalone accuracy. The pattern is quantified by a perfect rank correlation across architectures ($\rho = 1.0$, $p = 0.003$, $n{=}6$; Figure~\ref{fig:capability-gap}). On ProofWriter ($n{=}600$; Table~\ref{tab:cross_generator_nli}), only Mistral-7B (SC below DeBERTa's 55.00\%) gains significantly ($+6.67$\,pp, $p < 0.001$). At $w{=}5$, BART and RoBERTa \emph{degrade} Qwen2.5-14B ($-3.43$/$-4.41$\,pp; Appendix~\ref{app:multi-nli}).

\begin{table}[t]
\centering
\small
\caption{Cross-generator NLI escape with DeBERTa-large-mnli verifier (55.00\% standalone, ProofWriter $n{=}600$). Combo gain decreases monotonically with generator capability. Only Mistral-7B (SC below the verifier) shows a significant $w{=}3$ gain.}
\label{tab:cross_generator_nli}
\resizebox{\columnwidth}{!}{\begin{tabular}{@{}llcccc@{}}
\toprule
Generator & Family & SC & Gap & Combo $\Delta$ ($w{=}3$) & $p$ \\
\midrule
Mistral-7B & Mistral & 39.33\% & $+$15.67 & $+$6.67\,pp & $<$0.001 \\
LLaMA-3.1-8B & LLaMA & 72.50\% & $-$17.50 & $+$0.83\,pp & 0.665 \\
Qwen2.5-14B & Qwen & 74.00\% & $-$19.00 & $+$0.67\,pp & 0.503 \\
Qwen3-14B & Qwen & 85.83\% & $-$30.83 & $+$0.50\,pp & 0.453 \\
\bottomrule
\end{tabular}}
\end{table}

\paragraph{Domain-adapted verifier.}
\looseness=-1 Fine-tuning DeBERTa on ProofWriter depth-5 OWA (99.67\% standalone) amplifies escape: $+$12.67/$+$11.67\,pp for Mistral-7B/LLaMA-8B (up to $14\times$ over zero-shot (LLaMA-8B); Table~\ref{tab:finetuned-nli}), extending gains to generators with SC up to ${\sim}$72\%. A cost-aware cascade recovers 83--90\% of gain at ${\sim}$35\% cost (Appendix~\ref{app:drc}). These results serve as directional evidence (the domain-adapted verifier introduces training).

\paragraph{Cross-model voting vs.\ cross-architecture NLI.}
\looseness=-1 An ensemble diversity control replaces the NLI verifier with a second LLM as cross voter ($w{=}3$). With Mistral-7B as cross voter on ProofWriter, all three strong generators degrade significantly ($-2.33$ to $-9.33$\,pp, all $p \leq 0.020$); even a stronger cross voter (Qwen2.5-14B) degrades Qwen3-14B at $w{=}5$ ($-2.50$\,pp, $p = 0.006$). In contrast, DeBERTa NLI at comparable standalone accuracy (55.00\%) never degrades any generator ($\Delta \leq +0.83$\,pp). This asymmetry reflects the structural independence of encoder-only NLI---different architecture, training objective, and input representation---from autoregressive biases shared across LLM families (Table~\ref{tab:cross-model-ablation}).
